---
title: "ggplot "
output: github_document
---

I'm an R Markdown document! 
```{r}
library(tidyverse)
#package viridis
library(viridis)
library(patchwork)
library(ggridges)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp =.6,
  out.width = "90%"
)
```


# 10/5 Visualization 1

load in a dataset that we will use often
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Scatterplot
tmax vs tmin
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

you can save ggplot
```{r}
ggp_tmax_tin <- 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_tmax_tin
##ggsave()
```

## let's fancy it up

add color? lines? other stuff?
```{r}
## what in ggplot apply to everything
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,color = name )) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE)

## when aes in geom_point, it will only work for point

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax )) +
  ## we can use null to make it not apply
  ## geom_point(aes(color = NULL) ,alpha = 0.3)
  geom_point(aes(color = name) ,alpha = 0.3) +
  geom_smooth(se = FALSE)

## when aes in geom_point, it will only work for point
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name )) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(.~name)
```
let's make one more scatterplot
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax ,size = prcp)) +
  geom_point(alpha = .3) +
  facet_grid(.~name) +
  geom_smooth(se = FALSE)
```

## use data manipulation as part of this

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax = tmax * (9 / 5) + 32,
    tmin = tmin * (9 / 5) + 32
  ) %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

which geoms do you want
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name))+
  geom_smooth()

##hexbin and geom_hex which can be used for a alot of data
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))+
  geom_hex()

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))+
  geom_bin_2d()
```

## UNIvariate plots

```{r}
## fill = name to show the resource
weather_df %>% 
  ggplot(aes( x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(.~name)

```

Let's try some other plots...

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name))+
  geom_density(alpha = .3)
```

still with tmax and tmin
```{r}
weather_df %>% 
  ggplot(aes(x = name , y = tmax)) +
  geom_boxplot()
```

some people like violin pilot
```{r}
weather_df %>% 
  ggplot(aes(x = name , y = tmax , color =name)) +
  geom_violin()
```

What about rigdes...
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name))+
  geom_density_ridges(alpha =.8, scale = .8)
```

## Embedding plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3)
```

```{r, fig.width = 12,  fig.height = 5}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3)
```

# 10/7
## start with a familiar one

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name))+
  geom_point(alpha = .3)+
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temperature(c)",
    y = "Maximum daily temperature(c)",
    caption = "Data from the rnoaa package with three stations"
  )
```

## Scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temperature(c)",
    y = "Maximum daily temperature(c)",
    caption = "Data from the rnoaa package"
  )+
  scale_x_continuous(
    breaks = c(-15,0,15),
    labels = c("-15","0","15")
  ) +
  ##scale_y_continuous(
    ##breaks = c(10,20,30),
    ##labels = c("10","20","30")
  #)
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temperature(c)",
    y = "Maximum daily temperature(c)",
    caption = "Data from the rnoaa package"
  ) +
  scale_color_hue(
    name = "Location", 
    h = c(100, 300)) +
  #options can be other
  scale_color_viridis_d()
```

## Themes
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Minimum daily temperature(c)",
    y = "Maximum daily temperature(c)",
    caption = "Data from the rnoaa package with three stations"
  ) +
  scale_color_viridis_d() +
  # change the position
  # it is important to make sure the postion of theme command
  #theme_bw() + 
  #theme_classic() +
  theme(legend.position = "bottom")
```

## Setting for the whole file
```{r eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## `data` in geoms
```{r}
central_park <- 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

#point from waikiki and line from centralpark
waikiki %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park)
  
```

## `patchwork`
```{r}
ggp_tmax_tin <- 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name))+
  geom_point(alpha = .3)+
  geom_smooth()


ggp_prcp_dens <-
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .3)

tmax_date_p <- 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

## + for the same row ,and / for new row
ggp_tmax_tin / (ggp_prcp_dens + tmax_date_p)
```

## data manipulation
quick example on factors
```{r}
weather_df %>% 
  mutate(name = fct_reorder(name, tmax)
  ) %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()

```

What about tmax and tmin 。...
```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "obs",
    values_to = "temperature"
  ) %>% 
  ggplot(aes(x = temperature, fill = obs)) +
  geom_density(alpha = .3) +
  facet_grid(.~name)
```

```{r}
pulse_df <- 
  haven::read_sas("data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi",
    names_prefix = "bdi_score_"
  ) %>% 
  mutate(
    visit = recode(visit, "bl" = "00m")
    )

pulse_df %>% 
  ggplot(aes(x = visit, y = bdi)) +
  geom_boxplot()

pulse_df %>% 
  ggplot(aes(x = visit, y = bdi)) +
  geom_point(size = .2) +
  geom_line(aes(group = id),alpha = .3)
```

#10.12 EDA1
```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    #use lubridate the day of month to a month
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

## `group_by`
```{r}
weather_df %>% 
  group_by(name,month)
```

## count some thing
```{r}
# for the further work
weather_df %>% 
  group_by(month) %>% 
  summarize(n_obs = n())

#jusk a quick way
weather_df %>% 
  count(name)
```

it is nice tah this is a data frame
```{r}
#it will produce a table, which can not be used in ggplot2
weather_df %>% 
  pull(name) %>% 
  table()
```

```{r}
#create a dataframe
weather_df %>% 
  janitor::tabyl(month,name)
```

## more general summaries

Let's count, take means, look at SDs

```{r}
weather_df %>% 
  group_by(name,month) %>% 
  summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmax = median(tmax, na.rm = TRUE),
    sd_tmax = sd(tmax, na.rm = TRUE)
  ) %>% 
  #can be used in ggplot
  ggplot(aes(x = month, y = mean_tmax, color = name))+
  geom_point()+
  geom_line()
```

fomatting table outputs
```{r}
weather_df %>% 
  group_by(name) %>% 
   summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = TRUE)
   ) %>% 
  #digits for the number
  knitr::kable(digits = 2)
```

## grouped `mutate`
```{r}
weather_df %>% 
  # with or without group are different
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    center_tmax = tmax - mean_tmax
  ) %>% 
  ggplot(aes(x = date, y = center_tmax , color = name))+
  geom_point()

```

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    tmax_rank = min_rank(tmax),
    #warmest:tmax_rank = min_rank(desc(tmax)
  ) %>% 
  filter(tmax_rank < 2)
```


lagged variable
```{r}
#lag 滞后函数
weather_df %>% 
  group_by(name) %>% 
  mutate(
    lagged.tmax = lag(tmax, n = 1),
    tmax_diff = tmax - lagged.tmax
  ) %>% 
  summarize(diff_sd = sd(tmax_diff,na.rm = TRUE))
```


## limitations

what is my "summary" is a linear model ...
```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(cor_tmin_max = cor(tmin,tmax,use = "complete"))

weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  lm(tmax ~ tmin,data = .)

```





