---
title: "ggplot "
output: github_document
---

I'm an R Markdown document! 
```{r}
library(tidyverse)
library(ggridges)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp =.6,
  out.width = "90%"
)
```


# 10/5 Visualization 1

load in a dataset that we will use often
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```
## Scatterplot
tmax vs tmin
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

you can save ggplot
```{r}
ggp_tmax_tin <- 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_tmax_tin
##ggsave()
```

## let's fancy it up

add color? lines? other stuff?
```{r}
## what in ggplot apply to everything
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,color = name )) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE)

## when aes in geom_point, it will only work for point

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax )) +
  ## we can use null to make it not apply
  ## geom_point(aes(color = NULL) ,alpha = 0.3)
  geom_point(aes(color = name) ,alpha = 0.3) +
  geom_smooth(se = FALSE)

## when aes in geom_point, it will only work for point
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name )) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(.~name)
```
let's make one more scatterplot
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax ,size = prcp)) +
  geom_point(alpha = .3) +
  facet_grid(.~name) +
  geom_smooth(se = FALSE)
```

## use data manipulation as part of this

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax = tmax * (9 / 5) + 32,
    tmin = tmin * (9 / 5) + 32
  ) %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

which geoms do you want
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name))+
  geom_smooth()

##hexbin and geom_hex which can be used for a alot of data
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))+
  geom_hex()

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))+
  geom_bin_2d()
```

## UNIvariate plots

```{r}
## fill = name to show the resource
weather_df %>% 
  ggplot(aes( x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(.~name)

```

Let's try some other plots...

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name))+
  geom_density(alpha = .3)
```

still with tmax and tmin
```{r}
weather_df %>% 
  ggplot(aes(x = name , y = tmax)) +
  geom_boxplot()
```

some people like violin pilot
```{r}
weather_df %>% 
  ggplot(aes(x = name , y = tmax , color =name)) +
  geom_violin()
```

What about rigdes...
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name))+
  geom_density_ridges(alpha =.8, scale = .8)
```

## Embedding plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3)
```

```{r, fig.width = 12,  fig.height = 5}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3)
```





